---
title: "FML_Assignment_1"
author: "Aditya Ashish Kulkarni"
date: "2024-02-01"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)

```


**Answer 1:** downloaded data from Kaggle link: <https://www.kaggle.com/datasets/sujay1844/used-car-prices>

**Answer 2:** Imported data set into R 

```{r}
df = read.csv('/Users/aditya/Downloads/train.csv')
head(df)

```

**Answer 3:** Print out of descriptive statistics 

``` {r}
summary(df)


``` 

``` {r}
str(df)

```

**Answer 4:** Transformation of variable. 

Normalization in range 0 - 1 

Also using pre process function for normalization
Getting all numeric values in the range between 0-1 

```{r}

#Preprocess definition for scaling to 0-1 range
preproc <- preProcess(df, method = c("range"))

# Apply normalization
normalized_data <- as.data.frame(predict(preproc, df))

# Print normalized data
head(normalized_data)

```



**Answer 5:**

Plotting data, 

```{r}
plot(df$Price, xlab = "Count of Cars", ylab = "Price of Cars", col = "blue")

```


Plotting seats:
Using table function for segregate seats and their counts and plotting it into bar graph using bar plot. 
````{r}
t = table(df$Seats)
barplot(t,xlab = "Number of Seats in Car", ylab = "Count", col = "cyan")
```




